---
title: "Ovary scRNA-Seq Exploration"
output:
    html_document:
        theme: united
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  cache.lazy = FALSE,
  tidy = TRUE
)
```

### Setup the Seurat Object
```{r init, message=FALSE}
library(Seurat)
library(dplyr)
library(Matrix)

# Load the ovary dataset
ovary.data <- Read10X(data.dir = "../../output/ovary1/outs/filtered_gene_bc_matrices/dm6.16")

# Load FBgn to Symbol map
fbgn.symbol <- read.table('/data/LCDB/lcdb-references/dmel/r6-11/gtf/dmel_r6-11.SYMBOL.csv', sep=',', header=TRUE, stringsAsFactors=FALSE)
colnames(fbgn.symbol) <- c('FBgn', 'symbol')
fbgn.symbol <- subset(fbgn.symbol, !is.na(fbgn.symbol$symbol))
```

```{r create.object, results='hide', message=FALSE}
# Initialize the Seurat object with the raw (non-normalized data).
# Keep all genes expressed in >= 3 cells (~0.1% of the data). Keep all cells with at least 200 detected genes
ovary <- CreateSeuratObject(raw.data = ovary.data, min.cells = 3, min.genes = 200, project = "10X_ovary")
nCells = dim(ovary@meta.data)[1]

par(mfrow=c(1, 2))
plot(density(ovary@meta.data$nGene), main='nGene')
plot(density(ovary@meta.data$nUMI), main='nUMI')
```

***

### QC and selecting cells for further analysis

```{r qc, results='hide', fig.height=7,fig.width=13}
# The number of genes and UMIs (nGene and nUMI) are automatically calculated for every object by Seurat.
# For non-UMI data, nUMI represents the sum of the non-normalized values within a cell
# We calculate the percentage of mitochondrial genes here and store it in percent.mito using AddMetaData.
# We use object@raw.data since this represents non-transformed and non-log-normalized counts
# The % of UMI mapping to MT-genes is a common scRNA-seq QC metric.
# NOTE: You must have the Matrix package loaded to calculate the percent.mito values.
# mito.genes <- grep(pattern = "^MT-", x = rownames(x = ovary@data), value = TRUE)
# percent.mito <- colSums(ovary@raw.data[mito.genes, ]) / colSums(ovary@raw.data)

# AddMetaData adds columns to object@data.info, and is a great place to stash QC stats
# ovary <- AddMetaData(object = ovary, metadata = percent.mito, col.name = "percent.mito")
VlnPlot(object = ovary, features.plot = c("nGene", "nUMI"), nCol = 2)
```

Looking at the violin plots, we have a large number of cells with between 2,000
and 4,000 expressed genes. The Seurat user guide uses a cutoff of
`200>x>=2,500`, but I think this would remove too much data from this sample. I
am going to change the cutoff to be `200>x>=4,500` to remove the extreme cells.

```{r qc_filter, results='hide', fig.height=7,fig.width=13}
# GenePlot is typically used to visualize gene-gene relationships, but can be used for anything
# calculated by the object, i.e. columns in object@data.info, PC scores etc.
# Since there is a rare subset of cells with an outlier level of high mitochondrial percentage
# and also low UMI content, we filter these as well
#par(mfrow = c(1, 2))
#GenePlot(object = ovary, gene1 = "nUMI", gene2 = "percent.mito")
GenePlot(object = ovary, gene1 = "nUMI", gene2 = "nGene")

# We filter out cells that have unique gene counts over 2,500 or less than 200
# Note that low.thresholds and high.thresholds are used to define a 'gate'
# -Inf and Inf should be used if you don't want a lower or upper threshold.
ovary <- FilterCells(object = ovary, subset.names = c("nGene"), low.thresholds = c(200), high.thresholds = c(4500))
nCells_filter = dim(ovary@meta.data)[1]
```

***

### Normalizing the data
I kept the settings in Seurat manual. The only parameter we may want to change
is the 10,000 scaling factor, which I think is supposed to approximate the
number of nUMI per-cell.

```{r normalize, results='hide'}
ovary <- NormalizeData(object = ovary, normalization.method = "LogNormalize", scale.factor = 1e4)
```

### Detection of variable genes across the single cells
I kept the dispersion cutoffs the same as the Seurat manual. The cutoffs may need a little adjusting.


```{r var_genes, fig.height=7, fig.width=11, results='hide'}
ovary <- FindVariableGenes(object = ovary, mean.function = ExpMean, dispersion.function = LogVMR, x.low.cutoff = 0.0125, x.high.cutoff = 3, y.cutoff = 0.5)
```

```{r len_var_genes}
length(x = ovary@var.genes)
```

***

### Scaling the data and removing unwanted sources of variation
Here I regress out the variability caused by the nUMI. We can think of this as
the block effect of having different amounts of expression per cell. There may
be other effects that we want to regress out (i.e., rRNA, mito, ...).

```{r regress, fig.height=7, fig.width=11, results='hide'}
ovary <- ScaleData(object = ovary, vars.to.regress = c("nUMI"))
```

***

### Perform linear dimensional reduction
```{r pca}
ovary <- RunPCA(object = ovary, pc.genes = ovary@var.genes, do.print = TRUE, pcs.print = 1:5, genes.print = 5, pcs.compute=100)
```

```{r pca_viz}
# Examine and visualize PCA results a few different ways
#PrintPCA(object = ovary, pcs.print = 1:5, genes.print = 5, use.full = FALSE)
VizPCA(object = ovary, pcs.use = 1:2)
PCAPlot(object = ovary, dim.1 = 1, dim.2 = 2)

# ProjectPCA scores each gene in the dataset (including genes not included in the PCA) based on their correlation
# with the calculated components. Though we don't use this further here, it can be used to identify markers that
# are strongly correlated with cellular heterogeneity, but may not have passed through variable gene selection.
# The results of the projected PCA can be explored by setting use.full=T in the functions above
ovary <- ProjectPCA(object = ovary, do.print = FALSE)
```

```{r multi-heatmap, fig.height=12, fig.width=9, warning=FALSE}
PCHeatmap(object = ovary, pc.use = 1:12, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, use.full = FALSE)
```

***

### Determine statistically significant principal components
Looking at all of these plots I think it is safe to use 30 PCs, but it may be
necessary to step out a bit more.

```{r elbow_plot, fig.height=6, fig.width=10, warning=FALSE}
PCElbowPlot(object = ovary, num.pc=100)
```

```{r jackstraw, fig.height=6, fig.width=10, warning=FALSE}
# NOTE: This process can take a long time for big datasets, comment out for expediency.
# More approximate techniques such as those implemented in PCElbowPlot() can be used to reduce computation time
ovary <- JackStraw(object = ovary, num.replicate = 100, do.print = FALSE, num.pc=30)
```

```{r jsplots, fig.height=20, fig.width=10, warning=FALSE}
JackStrawPlot(object = ovary, PCs = 1:30)
```

***

### Cluster the cells
Using PCs 1-17 generate clusters using KNN.

```{r cluster, fig.height=5, fig.width=7}
# save.SNN = T saves the SNN so that the clustering algorithm can be rerun using the same graph
# but with a different resolution value (see docs for full details)
ovary <- FindClusters(object = ovary, reduction.type = "pca", dims.use = 1:17, resolution = 0.6, print.output = 0, save.SNN = TRUE)
```

```{r cluster.params}
PrintFindClustersParams(object = ovary)
# While we do provide function-specific printing functions, the more general function to
# print calculation parameters is PrintCalcParams().
```

***

### Run Non-linear dimensional reduction (tSNE)

```{r tsne, fig.height=5, fig.width=7, message=FALSE, warnings=FALSE}
ovary <- RunTSNE(object = ovary, dims.use = 1:17, do.fast = TRUE)
```

```{r tsneplot, fig.height=5, fig.width=7}
# note that you can set do.label=T to help label individual clusters
TSNEPlot(object = ovary)
```

```{r saveobject}
save(ovary, file = "../../output/ovary_seurat.Robj")
```

***
### Finding differentially expressed genes (cluster biomarkers)

```{r markers1, fig.height=8, fig.width=15, warning=FALSE}
# find markers for every cluster compared to all remaining cells, report only the positive ones
ovary.markers <- FindAllMarkers(object = ovary, only.pos = TRUE, min.pct = 0.25, thresh.use = 0.25)
ovary.markers %>% group_by(cluster) %>% top_n(2, avg_diff)
merged = left_join(ovary.markers, fbgn.symbol, by = c('gene' = 'FBgn'))
write.table(merged[, c('gene', 'symbol', 'cluster', 'p_val', 'avg_diff', 'pct.1', 'pct.2')], file="../../output/ovary_preliminary_analysis_biomarkers.tsv", quote=FALSE, sep="\t", row.names=FALSE, col.names=TRUE)
```

NOTE: The tutorial has some additional downstream things, but not doing to do them for now.
